---
title: "Spotify songs"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny 
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(DT)
library(shiny)
library(tidyr)
library(tidyverse)
library(purrr)
library(here)
library(forcats)
```
```{r, message=FALSE}
# Check if the data already exist
if(!dir.exists(here("data"))) { dir.create(here("data")) }

if(!file.exists(here("data","tuesdata_song.RDS"))){
  tuesdata <- tidytuesdayR::tt_load('2020-01-21')
  spotify_songs <- tuesdata$spotify_songs
    
  # save the files to RDS objects
  saveRDS(tuesdata$spotify_songs, file = here("data","tuesdata_song.RDS"))
}
```

```{r, message=FALSE}
song<- readRDS(here("data","tuesdata_song.RDS"))
```

```{r}
songs_era <- song %>% 
  mutate(
    track_album_release_date = as.Date(track_album_release_date),
    year = as.numeric(format(track_album_release_date,'%Y')),
    era = case_when(year <= 2000 ~ "Before 2000",
                    year > 2000 & year <= 2010 ~ "2000-2010",
                    year > 2010 ~ "After 2010"),
    duration_min = duration_ms/60000)
```

About {data-icon="fa-question-circle"}
===================================== 

Column {data-width="70"}
-------------------------------------

**Purpose:** We want to show the characteristics of songs in relation to the creater and time of creation.

**Data:** The [data](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-01-21) includes the album, artist, playlist, popularity and a set of indexes to measure the musical characteristics of the song.

**Data Dictionary:** Here is a list of full [data dictionary](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-01-21#readme) for what all the column names mean.

The Data {data-icon="fa-database"}
===================================== 

Column {data-width=650}
-----------------------------------------------------------------------
### The Data 
The [data](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-01-21) includes the album, artist, playlist, popularity and a set of indexes to measure the musical characteristics of the song.

```{r}
songs_era <- song %>% 
  mutate(
    track_album_release_date = as.Date(track_album_release_date),
    year = as.numeric(format(track_album_release_date,'%Y')),
    era = case_when(year <= 2000 ~ "Before 2000",
                    year > 2000 & year <= 2010 ~ "2000-2010",
                    year > 2010 ~ "After 2010"),
    duration_min = duration_ms/60000)
```

```{r, echo=FALSE}
DT::renderDataTable({
  DT::datatable(songs_era,
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '450px'))
})
```

Static: Time of release 
===================================== 
Column {data-width=700 .tabset .tabset-fade}
-----------------------------------------------------------------------

### Number of songs with time
```{r}
songs_era %>%
  filter(year>2000) %>%
  group_by(year) %>% 
  mutate(
  Nsongs= n()) %>%
  ggplot(aes(fill=playlist_genre,x=year))+geom_histogram(aes(y = Nsongs),stat = "identity")+
  labs(title ="Number of songs released in each year after 2000 and the genre", subtitle = 
         "Spotify songs", caption = "Tingxi Long")+ylab("Number of songs")+xlab("Year of release")
```

### Popularity of songs with time
```{r}
songs_era %>%
  filter(year>2000) %>%
  group_by(year) %>% 
  mutate(
  popularity= mean(track_popularity)) %>%
  ggplot(aes(x=year,y=popularity))+geom_point()+geom_line()+
  labs(title ="Average popularity score by year", subtitle = 
         "Spotify songs", caption = "Tingxi Long")+ylab("Average popularity")+xlab("Year of release")
```



Static: Artists 
===================================== 


```{r}
df1<- songs_era%>%
  dplyr::count(track_artist)

df1 %>%
  arrange(desc(n)) %>%
  head(30) %>%
  ggplot()+geom_bar(aes(x=n, y=fct_reorder(.f=track_artist, .x = n)),stat = "identity")+
  labs(title ="Top 30 artist with highest number of songs", subtitle = 
         "Spotify songs", caption = "Tingxi Long")+ylab("Artist")+xlab("Number of tracks")
```

Explore genre {data-icon="fa-flag"}
===================================== 



Column {.sidebar}
-----------------------------------------------------------------------

To see how popularity varies across playlist genres.

```{r}
selectInput("genre", "Genre:",
                  c("EDM (edm)" = "edm",
                    "Latin (latin)" = "latin",
                    "Pop music (pop)" = "pop",
                    "R & B (r&b)" = "r&b", 
                    "Rap music (rap)" = "rap",
                    "Rock music (rock)"= "rock"))
```


Column {data-width=700 .tabset .tabset-fade}
-----------------------------------------------------------------------
### Popularity over genre

```{r}
formulaText <- reactive({
    paste("Popularity ~", input$genre)
  })

  
renderPlot({
    songs_era %>%
    filter(playlist_genre== input$genre) %>%
    filter(track_popularity != 0) %>%
    group_by(track_popularity)%>%
    mutate(
      N= n()
    )%>%
    ggplot(aes(fill=year,x=track_popularity,y=N))+
    geom_bar(stat='identity')+ 
    labs(title ="Number of songs with populartiy under genre", caption = "Tingxi Long")+
    ylab("Number of songs")+xlab("Popularity")
  })

```
### dancebility over genre

```{r}
formulaText <- reactive({
    paste("Dancebility ~", input$genre)
  })


renderPlot({
  songs_era %>%
    filter(playlist_genre== input$genre) %>%
    ggplot(aes(x=danceability))+
    geom_histogram(aes(y=..density..),colour = 1,fill="pink")+  geom_density()+
    labs(title ="Number of songs with dancebility under genre", caption = "Tingxi Long")+
    ylab("Number of songs")+xlab("dancebility")
  })

```


Explore era
=====================================

Column {.sidebar}
-----------------------------------------------------------------------

To see how popularity varies across eras.
```{r}
selectInput("era", "Era:",
                  c("Before 2000"="Before 2000",
                    "2000-2010"="2000-2010",
                    "After 2010"="After 2010"))
```

Column 
-----------------------------------------------------------------------
### Popularity over era

```{r}
formulaText <- reactive({
    paste("Popularity ~", input$era)
  })

renderPlot({
    songs_era %>%
    filter(era == input$era) %>%
    filter(track_popularity != 0) %>%
    group_by(track_popularity)%>%
    mutate(
      N= n()
    )%>%
    ggplot(aes(fill=playlist_genre,x=track_popularity,y=N))+
    geom_bar(stat='identity')+ 
    labs(title ="Number of songs with populartiy under different era", caption = "Tingxi Long")+
    ylab("Number of songs")+xlab("Popularity")
  })

```


Analysis
=====================================
## Era
### Popularity by Era
```{r}
f <- factor(c("2000", "00-10", "2010"), levels = c("2000", "00-10", "2010"))
songs_era %>%
  filter(!is.na(era)) %>%
  group_by(playlist_genre, era) %>%
  ggplot() +
  geom_boxplot(aes(x=fct_inorder(era), track_popularity, fill=era), outlier.size=0) +
  facet_wrap(~playlist_genre,nrow = 3) + 
  labs(x='Era',y="popularity") +
  theme_minimal() +
  theme(legend.position = 'none')+
  labs(title = "Track Popularity by Era and Playlist Genre",
       subtitle = "Data from Spotify", caption = "Tingxi Long",
       x='Era',
       y="popularity")
```


### Song duration
```{r}
songs_era %>%
  na.omit(era) %>%
  ggplot(aes(x=playlist_genre,y=duration_min)) +
  geom_violin(aes(fill=playlist_genre)) +
  facet_grid(~fct_inorder(era)) +
  theme_minimal() +
  labs(x='Duration (min)') +
  labs(title = "Song Duration by Playlist Genre and Era",
       subtitle = "Data from Spotify", caption = "Tingxi Long",
       x='Duration (min)') +
  theme(legend.position = 'none')
```

## Column2
### Loudness
```{r}
fig2 <-song %>%
  ggplot(aes(fill=playlist_genre,x=energy,y=loudness))+
  geom_bar(stat='identity')+ 
  labs(title ="Loudness of songs in relation to its energy index", subtitle = 
         "EDM has the largest loudness", caption = "Tingxi Long")+ylab("Loudness")+xlab("Energy index")
fig2
```
